Board Game Recommendation

### 2015-03-31 paper.Rnw
couple of comments

\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^n \left(y_i - \widehat{y_i}\right)^2
\end{equation}

@@

Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers. Then:

@@

The weights for this will be $\frac{1}{1 + \rho_{u,k}}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$. The number of neighbors to use $k$ will be determined by cross-validation on a user-to-user basis. In the end $z_{u,i}$ will be the k-nearest-neighbors estimate for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$:

\begin{equation}
z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} * \frac{1}{1 + \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 + \rho_{u,u_{[j]}}}}
\end{equation}

\subsubsection{Back to Rating}
From what we've produced above, of the estimated rating for user $u$ and game $i$ is given by:

@@

To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the parameter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation (1)).
RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s). Since the game parameter is fit first, we fit the $\alpha$ tuning parameter for the game effect first. After finding an optimal value, we'll then fit the $\alpha$ tuning parameter for the player effect, as the game effect is needed to estimate the player effect.

The optimization will use ten-fold cross validation, wherein the data is randomly divided into 10 separate sections of roughly equal size. Then for each of the 10 sections, we'll treat the section as a test set and the other 9 to use as a training set. $\widehat{\theta}_g$ is fit first, and then from a grid of $\alpha_{game}$ values (where $\alpha_{game} \in [1,10000]$), a value of $\widehat{\theta}_i$ is produced for each of the $\alpha_{game}$ in the grid. For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated by:
\begin{equation}
\text{RMSE} = \frac{ \sum r_{u,i}^{(1)} - (\theta_g + \theta_i)}{n}
\end{equation}

where we sum over all rated games in the test set, and $n$ is the number of ratings in the test set. The value of $\alpha_{game}$ that has the lowest RMSE will be used. in figure ~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.

\begin{center}
\begin{figure}
\centering
<<game_alpha_optim, echo = F, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
load("game_graph2")
game_graph
@
\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values.}
\end{figure}
\end{center}

@@

\begin{figure}
<<k_optim, echo = F>>=
load("k_plot")
k_plot
@
\caption{\label{fig:k}}
\end{figure}

\section{Technical Implementation}
This algorithm is implemented in R using shiny as a vehicle for a simple, easy to use user interface. Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratings if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user. This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the information to the user in a format which they will understand.

In an implementation of this, the application will begin by having the user enter the username they've registered on \url{boardgamegeek.com}. As seen in (~\ref{fig:blanksc}), there is an area to enter the users information in, and a button just below that, when clicked, begins the process.
\begin{center}
\begin{figure}
\includegraphics[scale=0.4]{blank_login.png}
\caption{\label{fig:blanksc} A blank log in screen}
\end{figure}
\end{center}
Instead of logging the users ratings directly through the application, we will instead leverage the existing infrastructure of boardgame geek. After checking that its a valid username, the application will scour boardgamegeek to see if the database's information is complete: that is, do we have this user's data in the player key? Have they rated games that aren't in our database? Have they re rated games that are in our database? Have they rated games that aren't in our database? If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game). After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we'll proceed to model fitting.

@@

Finally, after the 10-fold cross-validation, we're ready to run the other games through the algorithm to get a predicted score. The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user in the "My Recommendations" Tab, which can be seen in Figure ~\ref{fig:act_use}.
\begin{figure}
\includegraphics[scale=0.4]{active_use.png}
\caption{\label{fig:act_use} The "My Recommendations" tab, populated for user cwmassey.}
\end{figure}

@@

\begin{figure}
<<subdom_plot, echo = FALSE, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
load("classifiers_plot")
out
@
\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.}
\end{figure}

\begin{figure}
\includegraphics[scale=0.4]{cat_ratings.png}
\caption{\label{fig:cat_rat_tab} The "My Category Ratings" tab.}
\end{figure}

>>>>

\begin{equation}\label{eq:rmse}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^n \left(y_i - \widehat{y_i}\right)^2}
\end{equation}

@@

Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers. Then: \hh{XXX something went wrong with the formula ... not sure how to fix it}

@@

\hh{XXX why do those weights make sense? aren't those weights tilted towards players who are negatively correlated? are there any? (then it's not  a practical problem, even though still a theoretical one) just check that it is not supposed to be $1 - \rho$ in the denominator. I'm pretty sure that's what's it supposed to be.}
The weights for this will be $(1 + \rho_{u,k})^{-1}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$. The number of neighbors to use $k$ will be determined by cross-validation on a user-to-user basis. In the end $z_{u,i}$ will be the \hh{weighted average of the} k-nearest-neighbors \hh{scores} for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$):

\begin{equation}
z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} \cdot \frac{1}{1 + \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 + \rho_{u,u_{[j]}}}}
\end{equation}

\subsubsection{Back to Rating}
From what we've produced above \hh{XXX to be more specific use the relevant equation numbers that are involved }, of the estimated rating for user $u$ and game $i$ is given by:

@@

To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the parameter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation \ref{eq:rmse}).
RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s). Since the game parameter is fit first, we fit the $\alpha$ tuning parameter for the game effect first. After finding an optimal value, we'll then fit the $\alpha$ tuning parameter for the player effect, as the game effect is needed to estimate the player effect.

The optimization will use ten-fold cross validation, wherein the data is randomly divided into 10 separate sections of roughly equal size. Then for each of the 10 sections, we'll treat the section as a test set and the other 9 to use as a training set. $\widehat{\theta}_g$ is fit first, and then from a grid of $\alpha_{game}$ values (where $\alpha_{game} \in [1,10000]$), a value of $\widehat{\theta}_i$ is produced for each of the $\alpha_{game}$ in the grid. For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated here by: \hh{XXX I am not sure this really is the RMSE - where did the squaring and the square rooting go?}
\begin{equation}
\text{RMSE} = \frac{ \sum r_{u,i}^{(1)} - (\theta_g + \theta_i)}{n}
\end{equation}

where we sum over all rated games in the test set, and $n$ is the number of ratings in the test set. The value of $\alpha_{game}$ that has the lowest RMSE will be used. in figure~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.

\begin{figure}
\centering
<<game_alpha_optim, echo = F, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
load("game_graph2")
game_graph
@
\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values.}
\end{figure}

@@

\begin{figure}\centering
<<k_optim, echo = F, out.width='0.45\\textwidth'>>=
load("k_plot")
k_plot
@
\caption{\label{fig:k} \hh{You still need a caption here.}}
\end{figure}

\section{Technical Implementation} \hh{give references for R (to be a nice R user), shiny, and the other packages that you are using. citation() and citation("packagename") makes your life easy with that.}
This algorithm is implemented in R  using shiny as a vehicle for a simple, easy to use user interface. Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratings if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user. This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the information to the user in a format which they will understand.

In an implementation of this, the application will begin by having the user enter the username they've registered on \url{boardgamegeek.com}. As seen in (~\ref{fig:blanksc}), there is an area to enter the users information in, and a button just below that, when clicked, begins the process.
\begin{center}
\begin{figure}\centering
\includegraphics[scale=0.4]{blank_login.png}
\caption{\label{fig:blanksc} A blank log in screen}
\end{figure}
\end{center}
Instead of logging the users ratings directly through the application, we will instead leverage the existing infrastructure of boardgame geek. After checking that its a valid username, the application will scour boardgamegeek to see if the database's information is complete: that is, do we have this user's data in the player key? Have they rated games that aren't in our database? Have they re rated games that are in our database? Have they rated games that aren't in our database? If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game). After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we'll proceed to model fitting.

@@

Finally, after the 10-fold cross-validation, we're ready to run the other games through the algorithm to get a predicted score. The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user in the "My Recommendations" Tab, which can be seen in Figure ~\ref{fig:act_use}.
\begin{figure}\centering
\includegraphics[scale=0.4]{active_use.png}
\caption{\label{fig:act_use} The "My Recommendations" tab, populated for user cwmassey.}
\end{figure}

@@

\begin{figure}\centering
<<subdom_plot, echo = FALSE, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
load("classifiers_plot")
out
@
\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.  \hh{XXX fix the y label, if showing it}}
\end{figure}

\begin{figure}\centering
\includegraphics[scale=0.4]{cat_ratings.png}
\caption{\label{fig:cat_rat_tab} The ``My Category Ratings" tab.}
\end{figure}
